Determinism, Complexity, and Predictability in Computer Performance

### 2014-05-30 pecomplexity.tex
fix an dash

This paper provides insight into when, why, and how forecast
strategies fail when they are applied to complicated time series. We
conjecture that the inherent complexity of a time series —-- which
results from the dimension, nonlinearity, and non-stationarity of the
generating process, as well as from measurement issues like noise,
aggregation, and finite data length --- is both empirically
quantifiable and directly correlated with predictability.  In
particular, we argue that \emph{redundancy} is an effective way to
measure complexity and predictive structure in a time series and that
\emph{weighted permutation entropy} is an effective way to estimate
that complexity.  To validate these conjectures, we study 120
different time-series data sets.  For each time series, we construct
predictions using a wide variety of forecast models, then compare the
accuracy of the predictions with the permutation entropy of that time
series.  We use the results to develop a heuristic that can help
practitioners recognize when a particular prediction method is not
well matched to the task at hand: that is, when the time series has
more predictive structure than that method can capture and exploit.

>>>>

This paper provides insight into when, why, and how forecast
strategies fail when they are applied to complicated time series. We
conjecture that the inherent complexity of a time series --- which
results from the dimension, nonlinearity, and non-stationarity of the
generating process, as well as from measurement issues like noise,
aggregation, and finite data length --- is both empirically
quantifiable and directly correlated with predictability.  In
particular, we argue that \emph{redundancy} is an effective way to
measure complexity and predictive structure in a time series and that
\emph{weighted permutation entropy} is an effective way to estimate
that complexity.  To validate these conjectures, we study 120
different time-series data sets.  For each time series, we construct
predictions using a wide variety of forecast models, then compare the
accuracy of the predictions with the permutation entropy of that time
series.  We use the results to develop a heuristic that can help
practitioners recognize when a particular prediction method is not
well matched to the task at hand: that is, when the time series has
more predictive structure than that method can capture and exploit.
